BACI of biomass change for VOCC source cells for all years

### 2019-08-29 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
optimal values integration with BACI code... incomplete

Should optimal temperature vary for age classes? Seems likely. 
Therefore need separate VOCC trimming and matching for age classes -> change function structure to choose age upfront

Should matches be selected based on exact depth categories in addition to density (all within time periods)?
What does it mean for the analysis that the depth range at which VOCC occur within a species temperature range varies dramatically between time periods?

@@

# climate <- "do"
# do_threshold <- c(0.1)

@@

@@

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, echo=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

if (age == "imm") {
  params <- filter(params, age == "imm")
} else {
  params <- filter(params, age != "imm")
}
params <- params[rev(order(params$maxy)), ]
params <- params[!duplicated(params$species), ]

#optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

#optimal_values <- c(optimal_do, optimal_temp)
```

@@

```{r, include=FALSE}
# Set default cell sizes and scale
input_cell_size <- 2
dist_intercept <- input_cell_size/2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  optimal_values <- optimal_temp
  max_thresholds <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  optimal_values <- optimal_do
  min_thresholds <- do_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")
  min_thresholds <- c(do_threshold, Inf)
  max_thresholds <- c(Inf, temp_threshold)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

# Calculate climate vectors

```{r}
rm(vocc)

@@

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- lapply(matched, function(i){  
  summary(i)
  plot(i)
  })
match_plots
}
```

@@

## Adults
```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp [["adult"]]
# View(data)
```

@@

```{r,  results="hide", include=FALSE}
plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source"), start_time=c("2009","2011","2013","2015")), parametricOnly = TRUE, main="", xlab ="log_density"
  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
  ) 
```

## Immatures
```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
```

### Plot matches
```{r}
idat_summary <- idata %>% 
  group_by(start_time, cell_type) %>%
  tally() %>% filter(cell_type=="source")
idata_depth <- idata %>% 
  group_by(start_time) %>% 
  mutate(mean_depth = mean(depth))

ggplot(idata, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
  geom_point(alpha=0.4, size = 1) +
  scale_shape_manual(values= c(0, 15)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()

lapply(variable_names, function(j) {
ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")+
  facet_grid(start_time~cell_type) +
  theme_bw() 
})

# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   geom_text(data = idata_depth,
#     aes(150, 0.5, label = round(mean_depth)), hjust = -0.3, inherit.aes = FALSE) +
#   theme_bw()
ggplot(idata, aes(depth, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~start_time) +
  theme_bw()

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(start_time~time) +
  geom_text(data = idat_summary,
            aes(1, 0.75, label = n), hjust = -0.3) +
  theme_bw()

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(time~start_time) +
  theme_bw()

ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  geom_text(data = idat_summary,
            aes(cell_type, max(data$log_density), label = n), hjust = -0.3) +
  facet_grid(time~start_time) +
  theme_bw()
```

### Basic BACI
```{r}
mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
summary(mi)
```

```{r}
sjPlot::plot_model(mi, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "immature"))
```

### GAMM
```{r}
# for (j in (variable_names)) {
#  gamm_formula <- paste0("log_density ~ ", 
#   # "time * start_time + ",
#   "start_time * cell_type + ",
#   "time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
#  }
igam4 <- gamm4::gamm4(as.formula(gamm_formula), 
  random = ~(1|matchobs) + (1|icell) 
   + (1|start_time)
  # + (time|start_time) 
  # + (cell_type|start_time)
  , data=idata)

summary(igam4$gam)

plot(igam4$gam,pages=1)
```

```{r, results="hide", include=FALSE}
plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), parametricOnly = TRUE, main="", xlab ="log_density"
  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
  ) 
```

Misc exploring gamm 
```{r eval=FALSE}
# summary(igam4$mer) ## underlying mixed model
# anova(igam4$gam)

# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), parametricOnly = TRUE, main="")

# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
# summary(mi)
# 
# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(vect_dist) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
# summary(mi)
# mi
# 
# 
# idata$after <- as.factor(idata$after)
# idata$matchobs <- as.factor(idata$matchobs)
# 
# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(vect_dist) + #s(X) + s(Y) + 
#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
# 
# 
# plot(m4b$gam,pages=1)
# 
# summary(m4b$gam) ## summary of gam
# summary(m4b$mer) ## underlying mixed model
# anova(m4b$gam) 
# 
# gamtabs(m1)
# 
# #library(itsadug)
# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
# 
# library(mgcv)
```

>>>>

Modified to allow optimal temperature to vary by maturity class. 
Therefore need to run this file separately for each species-class combination.

Should matches be selected based on exact depth categories in addition to density (all within time periods)?

What does it mean for the analysis that the depth range at which VOCC occur within a species temperature range varies dramatically between time periods?

@@

# climate <- "do"
# do_threshold <- c(0.5)

@@

#### Set default cell sizes and scale
```{r, echo=FALSE}
input_cell_size <- 2
dist_intercept <- input_cell_size/2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL
```

@@

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

if (age == "imm") {
  params <- filter(params, age == "imm")
} else {
  params <- filter(params, age != "imm")
}

params <- params[params$species == species, ]

if (climate == "temperature") {
  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  optimal_temps <- na.omit(optimal_temps)
  optimal_temp <- round(optimal_temps[1], digits = 2)
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  optimal_values <- optimal_temp
  max_thresholds <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  optimal_dos <- na.omit(optimal_dos)
  optimal_do <- round(optimal_dos[1], digits = 2)
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  optimal_values <- optimal_do
  min_thresholds <- do_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {
  optimal_temps <- params[rev(order(params$t_maxy)), ]$optimal_temp
  optimal_temps <- na.omit(optimal_temps)
  optimal_temp <- round(optimal_temps[1], digits = 2)
  optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  optimal_dos <- na.omit(optimal_dos)
  optimal_do <- round(optimal_dos[1], digits = 2)
  optimal_values <- c(optimal_do, optimal_temp)
  variable_names <- c("do_est", "temp")
  min_thresholds <- c(do_threshold, Inf)
  max_thresholds <- c(Inf, temp_threshold)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

@@

# Calculate climate vectors
```{r}
rm(vocc)

@@

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- apply(matched, function(matched){  
  summary(matched)
  plot(matched)
 })
match_plots
}
```

@@

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp [[1]]
# View(data)
```

@@

```{r, eval=FALSE, results="hide", include=FALSE}
plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source"), start_time=c("2009","2011","2013","2015")), parametricOnly = TRUE, main="", xlab ="log_density"
  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
  ) 
```
